---
title: "Data Aggregation and Formation of Final ADS"
author: "Team"
date: "10/28/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
# library(NbClust)
# library(factoextra)
library(plotly)

```


## Initial Solution

Aggregate the data and calculate the explanatory variables


```{r reading the datasets, echo=FALSE}
customerAttributes <- read.csv("Online_retail_Customer Detail/customer_data_2015_attr.csv")
print("The below dataset contains the demographic information of each customer like Age, Gender")
print(head(customerAttributes))

customerVisits <- read.csv("Online_retail_Customer Detail/customer_data_2015_visited.csv")
print("The below data shows how many each customer has visited the website")
print(head(customerVisits))

customerTransactions <- read.csv("Online_retail_Customer Detail/customer_data3_2015_transactions.csv")
print("The below data depicts the transactional information of each customer, spend per transaction and also the date of transaction")
print(head(customerTransactions))
```



## Merging the data
In order to obtain both the customer attributes and visits in a single dataset we do the process of merging. 
Prior to merging - need to check if the datasets for merging are at the same level or not

### Checking the level of datasets to be merged

```{r checking the level of datasets, echo=FALSE}
print("Checking the level of customerAttributes dataset..")
checkLevelAttr <- customerAttributes %>% group_by(CustomerId) %>% summarise(custPerRow = n())
levelAttr <- checkLevelAttr[which(checkLevelAttr$custPerRow > 1), ]
levelAttr <- data.frame(levelAttr)
if(nrow(levelAttr) == 0) {
  print("The customer attributes dataset is at customer level")
} else {
  print("The customer attributes dataset is not customer level")
}

print("Checking the level of customerVisits dataset..")
checkLevelVis <- customerVisits %>% group_by(CustomerID) %>% summarise(totalVisits = sum(Visits, na.rm = TRUE), custPerRow = n())
levelVis <- checkLevelVis[which(checkLevelVis$custPerRow > 1), ]
levelVis <- data.frame(levelVis)
if(nrow(levelVis) == 0) {
  print("The customer visits dataset is at customer level")
} else {
  print("The customer visits dataset is not customer level")
}

```




### Merging the visits and attributes datasets

```{r merging, echo=FALSE}

customerAttrVis <- merge(customerAttributes, customerVisits, by.x = "CustomerId", by.y = "CustomerID", all = FALSE)
checkLevelAttrVis <- customerAttrVis %>% group_by(CustomerId) %>% summarise(custPerRow = n())
levelAttrVis <- checkLevelAttrVis[which(checkLevelAttrVis$custPerRow > 1), ]
levelAttrVis <- data.frame(levelAttrVis)
if(nrow(levelVis) == 0) {
  print("The merged dataset is at customer level")
} else {
  print("The merged dataset is not customer level")
}
print(paste("The number of customers in the merged dataset:", nrow(customerAttrVis), sep = ""))
print(paste("The number of custoemrs in the attributes dataset:", nrow(customerAttributes), sep = ""))
print(paste("The number of customers in the visits dataset:", nrow(customerVisits), sep = ""))

if((nrow(customerAttributes) == nrow(customerVisits)) && (nrow(customerAttributes) == nrow(customerAttrVis))) {
  print("The merged dataset has the same number of customers as the two parent datasets")
}

```


## Analytical dataset creation for clustering

```{r variableTransformation, echo=FALSE}
### Transforming a few variables in the customerTransactions dataset
#Converting transaction date to a date column 
customerTransactions$transactionDate <- as.Date(customerTransactions$transactionDate, '%Y-%m-%d')
latestDate <- max(customerTransactions$transactionDate)
earliestDate <- min(customerTransactions$transactionDate)
dateRange <- dplyr::filter(customerTransactions, between(customerTransactions$transactionDate, earliestDate, latestDate))
if(nrow(dateRange) > 0) {
  print(paste("The transactions lie between: ", earliestDate, " to ", latestDate, sep = ""))
}

## Checking if any customer has more than one transaction in a day
sameDayTransact <- customerTransactions %>% group_by(CustomerID, transactionDate) %>% summarise(totalSpend = sum(Spends, na.rm = TRUE), no_of_tran = n_distinct(transactionID))

singleDayTran <- sameDayTransact[sameDayTransact$no_of_tran > 1, ]
distinctCustomers <- length(unique(singleDayTran$CustomerID))

print(paste("There are ", distinctCustomers, " customers who have transacted more than once in a day", sep = ""))
print(paste("Ignoring this insight as this doesn't really help me achieve anything"))

```

### Explanatory variables formation

```{r explanatory variables, echo=FALSE}
ADS <- customerTransactions %>% group_by(CustomerID) %>% summarise(totalSpends = sum(Spends), latestVisit = max(transactionDate), noOfTransactions = n_distinct(transactionID), recencyDays = (latestDate - max(transactionDate)))
ADS <- data.frame(ADS)

print(head(ADS))
adsNames <- colnames(ADS)

print(paste(adsNames[5], ": number of days between ", latestDate, " and the last day of the transaction of the customer", sep = ""))
print(paste(adsNames[4], ": Total number of transactions made by the customer within the year", sep = ""))
print(paste(adsNames[3], ": Date of the last transaction made by the customer", sep = ""))
print(paste(adsNames[2], ": Dollar value of all the transactions made by the customer", sep = ""))

## Merging back to the original dataset
adsFinal <- merge(customerAttrVis, ADS, by.x = "CustomerId", by.y = "CustomerID", all.x = TRUE)
# adsFinalBehaviour <- merge(customerAttrVis, ADS, by.x = "CustomerId", by.y = "CustomerID", all.y = TRUE)
print(head(adsFinal))
```






